# Import background basemap {#basemap}

```{r echo = FALSE, message = FALSE, warning=FALSE, results='hide'}
#load packages
libs <- c('ggplot2', "data.table", "dplyr", "cowplot", 
          "rnaturalearth", "rnaturalearthdata",
          "ggspatial", "sf")
lapply(libs, require, character.only = TRUE)

```

Use ```ne_states``` to get provincial/territorial borders from ```rnaturalearth``` package

```{r}
canada_map <- ne_states(country = 'canada', returnclass = "sf")
```

## Canada map and Newfoundland subset

- ```geom_sf``` function here allow to open the naturalearth map polygon. Inside this function, ```fill = ifelse ()``` allow to identify with a specific color the Newfoundland and Labrador region.
- ```coord_sf``` allow here to set the projection of our choice.
- ```scale_x_continuous``` allow to manually set the breaks of the x-axis.

```{r}
can_map <- ggplot() + 
  geom_sf(data = canada_map, colour = "black", 
          fill = ifelse(canada_map$name == "Newfoundland and Labrador", 
                        'orange', 'white')) + 
  coord_sf(crs = "+proj=lcc +lat_1=49 +lat_2=77 +lon_0=-91.52 +x_0=0 +y_0=0 
           +datum=NAD83 +units=m +no_defs") + 
  scale_x_continuous(breaks = c(-120, -100, -80, -60)) + 
  theme_bw() 

can_map
```


**Subset the map to Newfoundland part (need for next part of the manual):**

```{r}
map_nl <- canada_map[10,] 
```


