# Create interactive map with leaflet package {#interactive}

To learn about leaflet functions: (https://rstudio.github.io/leaflet/)

## Why interactive maps?: 

- Static maps are useful for creating figures for reports and presentation. Sometimes, however, you want to interact with your data. You can use the leaflet package for R to overlay your data on top of interactive maps. 
- Leaflet supports various elements for the composition of maps such as Map tiles, Markers, Polygons, Lines, Popups, and GeoJSON.
- Create a collaborative work, and share the results of your study, in an attractive way. 

```{r echo = FALSE, message = FALSE, warning=FALSE, results='hide'}
libs <- c('rnaturalearth', "rnaturalearthdata", "rgeos", "sf", "data.table", "ggplot2", "gganimate", "png", 'leaflet', 'dplyr')
lapply(libs, require, character.only = TRUE)
canada_map <- ne_states(country = 'canada', returnclass = "sf")
map_nl <- canada_map[10,] 

birdsurvey <- fread("Birdsurvey.csv")
Bubo <- subset(birdsurvey, species == 'Bubo scandiacus')
Bubo_2011_2021 <- subset(Bubo, year >= 2011 & year <= 2021)
```

## Add layers (markers, tiles, polygons)

```{r echo = TRUE, message = FALSE, warning=FALSE, results='hide'}
library(leaflet)
```

```{r, message = FALSE, warning=FALSE}
leaflet() %>%
  addTiles()
```

## Add the Snowy owl data

- ```addCircleMakers``` function to add the data with x (longitude) and y (latitude) axis. Similar than ```geom_point``` from ```ggplot```.


```{r}
# Other marker types, but circles are the simplest to use

leaflet() %>%
  addTiles() %>%
  addCircleMarkers(data = Bubo_2011_2021, lng = Bubo_2011_2021$decimalLongitude, lat = Bubo_2011_2021$decimalLatitude) 
```


## Customize Leaflet Maps

You can customize your leaflet map too. Let’s do the following:


* Add color palette by year with ```colorNumeric``` and ```colorFactor``` to set the palette (example of palettes:(https://www.datanovia.com/en/fr/blog/top-palettes-de-couleurs-r-a-connaitre-pour-une-meilleur-visualisation-des-donnees/)
* Adjust the point symbology.
* Add custom data-driven popups to your map.



```{r}
year <- colorNumeric("YlGnBu", domain = Bubo_2011_2021$year) 

year <- colorFactor(topo.colors(16), domain = Bubo_2011_2021$year) 

leaflet() %>%
  addTiles() %>%
  addCircleMarkers(data = Bubo_2011_2021, 
                   lng = Bubo_2011_2021$decimalLongitude, 
                   lat = Bubo_2011_2021$decimalLatitude,
                   fillColor = ~year(year), 
                   fillOpacity = 0.8,
                   color = "black",
                   weight = 1,
                   radius = 5) %>%
  addLegend(data = Bubo_2011_2021,
            pal = year,
            values = ~year,
            title = "Year")
```



## Add Scale bar

* To add a scale bar, need function ```addScaleBar```



```{r, warning=FALSE}
# ---- Mapping: Add Scale Bar ----
leaflet() %>%
  addTiles() %>%
  addCircleMarkers(data = Bubo_2011_2021, 
                   lng = Bubo_2011_2021$decimalLongitude, 
                   lat = Bubo_2011_2021$decimalLatitude,
                   fillColor = ~year(year), 
                   fillOpacity = 0.8,
                   color = "black",
                   weight = 1,
                   radius = 5) %>%
  addLegend(data = Bubo_2011_2021,
            pal = year,
            values = ~year,
            title = "Year") %>%
  addScaleBar(position = "bottomleft")
```

## Add Popups

* To add popups, you need to use **popup** in  ```addCircleMarkers``` and set the colum you want to show in your popup (here the localisation of the birds observed)


```{r, warning=FALSE}
# ---- Mapping: Add Popups ----
# Popups appear when you click on a point
leaflet() %>%
  addTiles() %>%
  addCircleMarkers(data = Bubo_2011_2021, 
                   lng = Bubo_2011_2021$decimalLongitude, 
                   lat = Bubo_2011_2021$decimalLatitude,
                   fillColor = ~year(year),  
                   fillOpacity = 0.8,
                   color = "black",
                   weight = 1,
                   radius = 5,
                   popup = ~locality) %>%
  addLegend(data = Bubo_2011_2021,
            pal = year,
            values = ~year,
            title = "Year") %>%
  addScaleBar(position = "bottomleft")
```


## Customize your popup 


* To customize your own popup, you need to use html
- <b>Hello</b>    b is **bold** tex Hello, in bold text
-  <br>    br adds a line break
* Then you can specify replace ```popup = ~locality``` with the name of your popup created (here **mypopup**)

```{r, warning = FALSE}
# ---- Mapping: Add Custom Popups ----
# Custom pop ups need to be written in html!
# <b>Hello</b>    b is bold tex Hello, in bold text
# <br>    br adds a line break
Bubo_2011_2021 <- Bubo_2011_2021 %>%
  mutate(mypopup = paste0(locality, "<br>",
                          "<b>Individual count: </b>", individualCount, "<br>",
                          "<b>Event date: </b>", eventDate, "<br>"))

# Now specify the popup as the custom popup we just created
leaflet() %>%
  addTiles() %>%
  addCircleMarkers(data = Bubo_2011_2021, 
                   lng = Bubo_2011_2021$decimalLongitude, 
                   lat = Bubo_2011_2021$decimalLatitude,
                   fillColor = ~year(year), 
                   fillOpacity = 0.8,
                   color = "black",
                   weight = 1,
                   radius = 5,
                   popup = ~mypopup) %>%
  addLegend(data = Bubo_2011_2021,
            pal = year,
            values = ~year,
            title = "Year") %>%
  addScaleBar(position = "bottomleft")
```



* If you want to save and share your collaborative map, you can use the function ```saveWidget``` from the package ```htmlWidgets```.

``` {r} 
#save it!
# htmlwidgets::saveWidget(map, "map.html", selfcontained = T)
```